---
title: "MVP"
output:
  html_document: default
  pdf_document: default
date: '2022-06-27'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, packages, results="hide", echo=FALSE, message = FALSE}
library(tidyverse)
library(tsibble)
library(fable)
library(GGally)
library(feasts)
library(knitr)
library(printr)
library(DT)
```
```{r, fake data, results="hide", echo=FALSE, message = FALSE}
# This is some time series data I found on the internet I transformed a bit
df <- read_csv("C:/Users/AxelTorbenson/OneDrive - eCapital Advisors, LLC/Documents/templates/CapitalForecasting/Data/working_capital_fake_data.csv")
```
```{r, tidying, results="hide", echo=FALSE}
# convert to tsibble and add working capital and current ratio
df$date <- as.Date(as.character(df$date), format = "%m/%d/%Y")
df <- df %>%
  mutate(date = yearmonth(as.character(date)),
         working_capital = assets - liabilities) %>%
  as_tsibble(index = date)

work_cap <- df %>%
  filter(row_number() <= n() - 36) %>%
  as_tsibble(index = date)
```

## Working Capital Forecasting Dashboard

This dashboard uses time series modeling to forecast future working capital levels.

This graph decomposes working capital levels into three parts: Trend, season, and remainder. The trend shows longer business cycles, the season shows seasonal fluctuations, and the remainder shows what is unaccounted for by the trend and seasonal components.

```{r, decomposition, echo=FALSE}
work_cap %>%
  model(STL(working_capital ~ trend(window = 21) +
              season(window = 13),
            robust = TRUE)) %>%
  components(trend) %>%
  autoplot() +
  ggtitle("Working Capital Decomposition") +
  labs(y = "Working Capital (millions)", x = "Date")
```

```{r, modeling, echo=FALSE}
arima_fit <- work_cap %>%
  model(# ARIMA function automatically picks best fitting model using AIC
    arima = ARIMA(working_capital, ic = 'aic', stepwise = FALSE),
    )
# refit to all data
full_fit <- arima_fit %>%
  refit(df)
```

The following graphs show three year forecasts for working capital levels. The first shows the forecast of working capital and the probability distribution around it. The second simulated five different paths working capital could take.

```{r, conf int graph, echo=FALSE}
forecast(full_fit, h = "3 year") %>%
  autoplot(filter(df, row_number() >= n() - 50), level = c(70, 95)) +
  ggtitle("Working Capital Forecast") +
  labs(y = "Working Capital (millions)", x = "Date")
```

```{r, sample paths, echo=FALSE}
simulation <- generate(full_fit, h = 10, times = 5) %>%
  filter(.model == 'arima')

df %>%
  filter(row_number() >= n() - 50) %>%
  autoplot(working_capital) + autolayer(simulation, .vars = .sim) +
  geom_line() +
  theme(legend.position = "none") +
  ggtitle("Working Capital Simulation") +
  labs(y = "Working Capital (millions)", x = "Date")
```

The following is a table that contains the forecasted working capital levels, as well as upper and lower bounds of the forecast. We are 95% sure working capital levels will be between the upper and lower bounds.

```{r, table, echo=FALSE}
three_yr_forecast <- forecast(full_fit, h = "3 year") %>%
  hilo(95) %>%
  as_tibble() %>%
  select(date, .mean, '95%') %>%
  rename(forecast = .mean) %>%
  rename(conf_int = '95%') %>%
  unpack_hilo(conf_int)

three_yr_forecast$conf_int_lower <-
  round(three_yr_forecast$conf_int_lower, 3)
three_yr_forecast$conf_int_upper <-
  round(three_yr_forecast$conf_int_upper, 3)
three_yr_forecast$forecast <- round(three_yr_forecast$forecast, 3)
three_yr_forecast$date <- as.Date(three_yr_forecast$date)

datatable(
  three_yr_forecast,
  class = 'cell-border stripe',
  colnames = c("Date", "Forecast", "Lower bound", "Upper bound"),
  caption = "Point forecast for working capital levels"
)
```
