---
title: "Price Elasticity Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    theme: bootstrap
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(shinyWidgets)
library(shinyjs)

# Core
library(tidyverse)
library(lubridate)

# Data Modeling
library(infer)
```


```{r import, include=FALSE}
product_summary_tbl <- read_rds("../data/product_summary.RDS")

product_analysis_tbl <- read_rds("../data/product_analysis.RDS")
```

Sidebar {.sidebar data-width=350}
===============================================================

```{r}
useShinyjs(rmd = TRUE)

shinyWidgets::pickerInput(
  inputId  = "picker_manu", 
  label    = h4("Manufacturer"), 
  choices  = unique(product_summary_tbl$MANU), 
  selected = c("Kellogg's"), 
  multiple = TRUE,
  options  = list(
    `actions-box` = TRUE,
    size = 10,
    `selected-text-format` = "count > 2"
  )
)

shinyWidgets::pickerInput(
  inputId  = "picker_product", 
  label    = h4("Product"), 
  choices  = unique(product_summary_tbl$DESCRIP), 
  selected = c("Frosted Flakes", "Rice Krispies", "Corn Flakes"), 
  multiple = TRUE,
  options  = list(
    `actions-box` = TRUE,
    size = 10,
    `selected-text-format` = "count > 2",
    `live-search` = TRUE
  )
)

shinyWidgets::pickerInput(
  inputId  = "picker_location", 
  label    = h4("Location"), 
  choices  = unique(product_summary_tbl$CITY), 
  selected = "Chicago", 
  multiple = TRUE,
  options  = list(
    `actions-box` = TRUE,
    size = 10,
    `selected-text-format` = "count > 3",
    `live-search` = TRUE
  )
)

shinyWidgets::pickerInput(
  inputId  = "picker_period", 
  label    = h4("Period"), 
  choices  = unique(product_summary_tbl$YEAR), 
  selected = unique(product_summary_tbl$YEAR), 
  multiple = TRUE,
  options  = list(
    `actions-box` = TRUE,
    size = 10,
    `selected-text-format` = "count > 3"
  )
)

br()
hr()
br()

actionButton(inputId = "apply", label = "Apply", icon = icon("play"))

actionButton(inputId = "reset", label = "Reset", icon = icon("sync"))

observeEvent(eventExpr = input$reset, handlerExpr = {
  
  updatePickerInput(
    session = session, 
    inputId = "picker_manu", 
    selected = c("Kellogg's")
  )
  
  updatePickerInput(
    session = session, 
    inputId = "picker_product", 
    selected = c("Frosted Flakes", "Rice Krispies", "Corn Flakes")
  )
  
  updatePickerInput(
    session = session,
    inputId = "picker_location",
    selected = "Chicago"
  )
  
  updatePickerInput(
    session = session,
    inputId = "picker_period",
    selected = unique(product_summary_tbl$YEAR)
  )
  
  shinyjs::delay(ms = 300, expr = {
    shinyjs::click(id = "apply")
  })
 
  
})


manufacturer <- reactive({
  
  product_summary_tbl %>%
    filter(MANU %in% input$picker_manu)
  
})


observeEvent(manufacturer(), {
  
  choices <- unique(manufacturer()$DESCRIP)
    
  updatePickerInput(
    session = session, 
    inputId = "picker_product", 
    choices = choices
  ) 
  
})

br()
hr()
br()
```

#### Definitions

**Elasticity** - Percentage change in Sales following a 1% change in Price.

**Inelastic** - Sales change *less* than 1% following a 1% change in Price. (Price Increase -> Revenue Gain)

**Elastic** - Sales change *more* than 1% following a 1% change in Price. (Price Increase -> Revenue Loss)

**Example**  

A 1% increase in ASP from $3.13 to $3.16 will decrease sales between 1.54% and 1.50%. This results in an expected revenue loss.



```{r filter}
product_summary_filtered_tbl <- eventReactive(
  eventExpr = input$apply, 
                                             
  valueExpr = {
  
    product_summary_tbl %>%
      
      filter(DESCRIP %in% input$picker_product) %>%
      
      filter(CITY %in% input$picker_location) %>%
      
      filter(YEAR %in% input$picker_period)
    
  },
  ignoreNULL = FALSE
)



product_analysis_filtered_tbl <- eventReactive(
  eventExpr = input$apply, 
                                             
  valueExpr = {
  
    product_analysis_tbl %>%
      
      filter(DESCRIP %in% input$picker_product) %>%
      
      filter(CITY %in% input$picker_location) %>%
      
      filter(YEAR %in% input$picker_period)
    
  },
  ignoreNULL = FALSE
)
```

```{r model}
product_elasticity_slopes_tbl <- reactive({

  product_analysis_filtered_tbl() %>%
    specify(formula = LOG_QUANTITY ~ LOG_PRICE) %>%
    generate(reps = 100, type = "bootstrap") %>%
    calculate(stat = "slope")

})

product_elasticity_ci_tbl <- reactive({

  product_elasticity_slopes_tbl() %>%
    get_confidence_interval(type = "percentile", level = 0.95)

})
```

Sales Overview
===============================================================

Row {data-height=150}
---------------------------------------------------------------

```{r}
kpi_tbl <- reactive({
  
  product_summary_filtered_tbl() %>%
    summarize(
      UNIT_SALES = sum(UNIT_SALES),
      REVENUE = sum(REVENUE),
      GROSS_MARGIN = sum(GROSS_MARGIN),
      GROSS_MARGIN_PCT = round(sum(GROSS_MARGIN) / sum(UNIT_SALES), 2)
    )
  
})
```


### Unit Sales

```{r}
renderValueBox({

  valueBox(
    value   = kpi_tbl()$UNIT_SALES %>% scales::comma(),
    caption = "Unit Sales"
  )

})
```


### Gross Revenue

```{r}
renderValueBox({
  
  valueBox(
    value   = kpi_tbl()$REVENUE %>% scales::dollar(), 
    caption = "Gross Revenue"
  )
  
})
```

### Gross Margin

```{r}
renderValueBox({
  
  valueBox(
    value   = kpi_tbl()$GROSS_MARGIN %>% scales::dollar(), 
    caption = "Gross Margin"
  )
  
})
```

### Gross Margin %

```{r}
renderValueBox({
  
  valueBox(
    value   = kpi_tbl()$GROSS_MARGIN_PCT %>% scales::percent(accuracy = 2), 
    caption = "Gross Margin %"
  )
  
})
```

Row {data-height=450}
---------------------------------------------------------------

### Revenue Contribution by Product

```{r}
output$plot_1 <- renderPlot({
  
  product_summary_filtered_tbl() %>%
    ggplot(aes(x = YEAR, y = REVENUE, fill = DESCRIP)) +
    geom_bar(position = "stack", stat = "identity") +
    ylab("Revenue") +
    xlab("Year") +
    scale_x_discrete(limits = seq(2014, 2022)) +
    scale_y_continuous(labels = scales::dollar_format()) +
    scale_fill_discrete(name = "Product")
  
})

plotOutput(outputId = "plot_1")
```

### Revenue Trend by Product

```{r}
output$plot_2 <- renderPlot({
  
  product_summary_filtered_tbl() %>%
    ggplot(aes(x = YEAR, y = REVENUE, group = DESCRIP, color = DESCRIP)) +
    geom_line() +
    ylab("Revenue") +
    xlab("Year") +
    scale_x_discrete(limits = seq(2014, 2022)) +
    scale_y_continuous(labels = scales::dollar_format()) +
    scale_color_discrete("Product")
  
})

plotOutput(outputId = "plot_2")
```


Row {data-height=450}
---------------------------------------------------------------

```{r}
product_summary_wide_tbl <- reactive({
  
  product_summary_filtered_tbl() %>%
    pivot_longer(c(UNIT_SALES:GROSS_MARGIN_PCT),
                 names_to = "METRIC",
                 values_to = "VALUE") %>%
    pivot_wider(names_from = YEAR, values_from = VALUE) %>%
    rename(Manufacturer = MANU, Product = DESCRIP, Location = CITY, Indicator = METRIC) %>%
    mutate(
      Indicator = recode(
        Indicator,
        UNIT_SALES = "Unit Sales",
        UNIT_SALES_GROWTH = "Unit Sales Y/Y Growth",
        AVG_RETAIL_PRICE = "Average Selling Price (ASP)",
        BUNDLE_SALES = "Bundle Sales",
        REVENUE = "Revenue",
        REVENUE_PCT = "Revenue %",
        GROSS_MARGIN = "Gross Margin",
        GROSS_MARGIN_PCT = "Gross Margin %"
      )
    )
  
})
```


### Product Detail

```{r}
output$table_1 <- renderTable({
  
  product_summary_wide_tbl()
  
}, hover = TRUE)

tableOutput(outputId = "table_1")
```


Statistical Analysis
===============================================================

```{r}
asp_tbl <- reactive({
  
  product_analysis_filtered_tbl() %>%
    summarize(
      AVG_RETAIL_PRICE = mean(exp(LOG_PRICE))
    )
  
})

elasticity_label_tbl <- reactive({

  product_elasticity_ci_tbl() %>%
    mutate(
      label = case_when(
        (lower_ci < -1) & (upper_ci < -1) ~ "Elastic",
        ((lower_ci > -1) & (upper_ci > -1)) & ((lower_ci < 0) & (upper_ci < 0)) ~ "Inelastic",
        TRUE ~ "Inconclusive"
      )
    )

})
```

Row {data-height=150}
---------------------------------------------------------------

### Average Unit Price

```{r}
renderValueBox({
  
  valueBox(
    value   = asp_tbl()$AVG_RETAIL_PRICE %>% scales::dollar(), 
    caption = "Average Selling Price (ASP)"
  )
  
})
```

### Elasticity Label

```{r}
renderValueBox({
  
  valueBox(
    value = elasticity_label_tbl()$label, 
    caption = "Elasticity Label"
  )
  
})
```

### Lower Bound

```{r}
renderValueBox({
  
  valueBox(
    value = round(product_elasticity_ci_tbl()$lower_ci, 2), 
    caption = "Lower Bound"
  )
  
})
```

### Upper Bound

```{r}
renderValueBox({
  
  valueBox(
    value = round(product_elasticity_ci_tbl()$upper_ci, 2), 
    caption = "Upper Bound"
  )
  
})
```

Row {data-height=450}
---------------------------------------------------------------
  
### Correlation Between Sales and Price
  
```{r}
output$plot_3 <- renderPlot({
  
  product_analysis_filtered_tbl() %>%
    ggplot(aes(x = LOG_PRICE, y = LOG_QUANTITY)) +
    geom_point(aes(color = DESCRIP), alpha = 0.5) +
    geom_smooth(se = FALSE, method = lm, color = "black") +
    xlab("Unit Price (Log Scale)") +
    ylab("Unit Sales / Bundle Sales (Log Scale)") +
    scale_x_continuous(labels = scales::dollar_format()) +
    scale_color_discrete(name = "Product")
  
})

plotOutput(outputId = "plot_3")
```

### Bootstrap Slopes Distribution

```{r}
output$plot_4 <- renderPlot({

  product_elasticity_slopes_tbl() %>%
    ggplot(aes(x = stat)) +
    geom_histogram() +
    geom_vline(xintercept = product_elasticity_ci_tbl()$lower_ci, linetype = "dashed", size = 1.25, color = "blue") +
    geom_vline(xintercept = product_elasticity_ci_tbl()$upper_ci, linetype = "dashed", size = 1.25, color = "blue") +
    xlab("Bootstrap Slopes (100 simulations)") +
    ylab("Frequency")

})

plotOutput(outputId = "plot_4")
```


Row {data-height=450}
---------------------------------------------------------------
  
### Product Detail
  
```{r}
output$table_2 <- renderTable({
  
  product_summary_wide_tbl()
  
}, hover = TRUE)

tableOutput(outputId = "table_2")
```