---
title: 'Elasticity Documentation'
author: "Lila Sahar and Juan Malaver"
date: "10/18/2021"
output: html_document
---

```{r, include = FALSE}
source("Elasticity_Functions.r", local = knitr::knit_global())
```

## Introduction

### What is Elasticity?

Elasticity is the change in consumers' demand in response to price changes. 

### Why is Elasticity important?

The elasticity helps businesses price their products and services, and it will 
help a company predict how the market will react when you do change a price of 
service or product.

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(readxl)

#importing data
descriptions_path <- "raw_data_cereal_descriptions.xlsx"
prices_path <- "raw_data_cereal_prices.xlsx"

descriptions_tbl <- input_descriptions(descriptions_path)
prices_tbl <- input_prices(prices_path)
```

## Business Problem

We will use *price elasticity of demand* ($\beta$) to understand how changes in
`price` impact changes in `sales`.

$$\text{log}(\texttt{sales})=\alpha+\beta\text{log}(\texttt{price})+\epsilon$$

## Cereal Data

```{r, echo = FALSE}
sales_tbl <- get_sales(descriptions_tbl, prices_tbl, top_three_brands_tbl)
```

## Distribution of Sales and Price by Brand

```{r, echo = FALSE}
plot_histogram_sales(sales_tbl, "Sales of Cereal Boxes", "Distribution of Sales")
plot_histogram_price(sales_tbl, "Price of Cereal Boxes", "Distribution of Prices")
```

## Cereal Data Sample

```{r, echo = FALSE}
sales_sample_tbl <- get_sales_sample(sales_tbl)
```

## Prices and Sales by Brand

```{r, echo = FALSE}
plot_scatter(sales_sample_tbl)
```

## Interpreting Price Elasticity
